Multi-task training of multi-domain NMT

### 2020-11-18 2021/mdmt/mdmt.tex
mdmt

>>>>

\section{Algorithm}
\begin{algorithm}[h]
\caption{Multi-domain Training}
\label{alg:multidomain}
\begin{algorithmic}[1]
\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Batch size $B$}%, Optimization algorithm $\operatorname{Opt}$}
\REPEAT 
\STATE{Iteration t}
\STATE{If $t mod 1000 == 0$: Update $H_t$ via \ref{alg:Fisher} else $H_t = H_{t-1}$}
\STATE{Randomly pick $i \in [1,..,d]$ w.r.t the multinomial distribution $[\frac{|C_i|}{\sum_{i\in [1,..,d]}|C_i|}]$.}
\STATE{Randomly pick $B$ sentences from $C_i$.}
\STATE{Activate residual adapters of domain $i$.}
\STATE{Calculate gradients $\frac{\delta L_i(B_i)}{\delta \theta_i}$, $\frac{\delta L_i(B_i)}{\delta \theta_s}$}
\STATE{Correct gradient by Fisher diagonal matrix $H_t$: 
\begin{align*}
\Delta \theta_i &= -\frac{1}{R} (H_t^i + \mu)^{-1} \frac{\delta L_i}{\delta \theta_i} \\
\Delta \theta_s &= -\frac{1}{R} (H_t^s + \mu)^{-1} \frac{\delta L_i}{\delta \theta_s} \\
R &=  (H_t^i + \mu)^{-1} \frac{\delta L_i}{\delta \theta_i} * \frac{\delta L_i}{\delta \theta_i}^T + (H_t^s + \mu)^{-1} \frac{\delta L_i}{\delta \theta_s} * \frac{\delta L_i}{\delta \theta_s}^T \\
\end{align*}
}
\STATE{Update parameters:
\begin{align*}
\theta_i^{t+1} = \theta_i^{t} - lr * \Delta \theta_i \\
\theta_s^{t+1} = \theta_s^{t} - lr * \Delta \theta_s \\
\end{align*}}
\UNTIL{convergence}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
\caption{Compute Fisher matrix}
\label{alg:Fisher}
\begin{algorithmic}[1]
\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Batch size $B$}
\STATE{Pick batch $B_i$ from $C_i$ $\forall i \in [1,..,d]$}
\FOR{$i\in [1,..,d]$}
     \STATE{$H^i = 0$, $H^s = 0$}
\ENDFOR
\FOR{$i\in [1,..,d]$}
   \FOR{$j\in [1,..,B]$}
        \STATE{Compute gradient wrt example $B_i^j$ of batch $B_i$: $\Delta H_i = \frac{\delta L_i(B_i^j)}{\delta \theta_i}$, $\Delta H_s =\frac{\delta L_i(B_i^j)}{\delta \theta_s}$}
        \STATE{Update $H_i$, $H_s$:
\begin{align*}
H_i &+= \Delta H_i ** 2 \\
H_s &+= \Delta H_s ** 2 \\
\end{align*}}
   \ENDFOR
\ENDFOR
\FOR{$i\in [1,..,d]$}
     \STATE{$H^i \setminus= B $, $H^s \setminus= B*d$}
\ENDFOR
\end{algorithmic}
\end{algorithm}
